##Dealing with Missing Data

```{r}
options(tidyverse.quiet = TRUE)

library(tidyverse)
library(tidymodels)
library(mice) #package for imputation
library(VIM) #visualizing missingness
library(naniar) #visualizing missingness
library(skimr) #alternative way to view dataset summaries
library(UpSetR) #visualizing missingness
```
```{r}
class_grades <- read_csv("class-grades.csv")
```
```{r}
grades=read_csv("class-grades.csv")
```



Read in dataset   
```{r}
grades=read_csv("class-grades.csv")
```

Structure and summary
```{r}
str(grades)
summary(grades)
skim(grades)
```

Visualizing missingness. There are MANY ways to look at missingness in R. Typically we are interested in proportion of missingness by variable, by case (row), and by factor.  

Here's a helpful link: https://cran.r-project.org/web/packages/naniar/vignettes/naniar-visualisation.html. NOTE: If you have lots of variables and/or rows, these plots can quickly become cluttered.  

Simple view of missingess  
```{r}
gg_miss_var(grades)
```

By case  
```{r}
gg_miss_case(grades) #x axis is number of missing values in each row (case)
```

By a factor (here we choose the variable Survived)  
```{r}
gg_miss_fct(x = titanic, fct = Survived)
```

Looking at missingness by variable and combinations of missingness using "aggr" from VIM package.  
```{r}
vim_plot = aggr(titanic, numbers = TRUE, prop = c(TRUE, FALSE),cex.axis=.7)
#the cex.axis reduces size of text on x-axis so labels fit better
```

A view of missingness by variable and row.  
```{r}
vis_miss(grades) #from the naniar package
```

So how do we do handle our missing data in this dataset? 

One idea: Column-wise deletion of the "Age" variable (creating a new data frame with this variable removed).  
```{r}
titanic_coldel = titanic %>% select(-Age) 
vim_plot = aggr(titanic_coldel, numbers = TRUE, prop = c(TRUE, FALSE),cex.axis=.7)
skim(titanic_coldel)
```

Second idea: Row-wise deletion of any row with at least one NA:
```{r}
grades_rowdel = grades %>% drop_na() 
#alternatively can specify which variable(s) on which to do the drop_na
#drop_na(Age)
skim(grades_rowdel)
```
```{r}
grades_coldel = grades %>% drop_na() 
#alternatively can specify which variable(s) on which to do the drop_na
#drop_na(Age)
skim(grades_coldel)
```
Doing this discards 177 rows of data.  

Finally, imputation. Let's use the "mice" package to do imputation.  
```{r}
set.seed(123) #sets seed for random number generator
imp_age = mice(grades, m=5, method='pmm', printFlag=FALSE)
#m is the number of imputations, 5 is a reasonable value as a default
#pmm is "predictive mean matching" = imputation method for numeric data
#printFlag reduces amount of output
summary(imp_age)
```
Merge the imputed values into our titanic data frame  
```{r}
grades_complete = complete(imp_age) 
summary(grades_complete)
```
